---
title: "chapter1"
author: "leuph1"
format: html
editor: visual
---

## Chapter1 - Generative Models for Discrete Data

```{r packages}
library(tidyverse)
library(cowplot)
library(khroma)
```

```{r}
three <- dpois(x = 3, lambda = 5)
dpois(x = 5, lambda = 5)
one <- dpois(x = 1, lambda = 5)
two <- dpois(x = 2 , lambda = 5)
zero <- dpois(x= 0, lambda = 5)

all <- one + two + three + zero
all
dpois(x = 5, lambda = 5)

dpois(x = 0:12, lambda = 5)
myPois <- dpois(x = 0:12, lambda = 5)
myX <- 0:12
myX
PoissonHIV <- as.data.frame(bind_cols(myX, myPois))
colnames(PoissonHIV) <- c("Mutations", "Probability")
head(PoissonHIV)
```

```{r}
PoissonHIV %>%
  mutate(color = "firebrick") %>%
  ggplot(aes(x = Mutations, y = Probability, fill = color)) +
  geom_col(position = position_dodge2()) +
  theme_cowplot() +
  scale_fill_identity() +
  scale_x_continuous(breaks = seq(0,12,1))
```

1.3 Discrete Probabilities

The outcomes of categorical variables (if outcomes == 2 =\> binary) are called levels.

Examples:

-   Coin toss: head or tail

-   Genotype: "AA", "Aa", "aa"

-   Number of different codons formed by 3 nucleotides: 4\*4\*4 = 64

Often we want to identify the frequency of outcomes in a vector of counts.

```{r}
genotype = c("AA","AO","BB","AO","OO","AO","AA","BO","BO",
             "AO","BB","AO","BO","AB","OO","AB","BB","AO","AO")
table(genotype)
```

```{r}
genotypeF <- factor(genotype)
levels(genotypeF)

```

We can also assign/reserve certain outcomes to be in the space of expected values, even if it is not observed (yet) in the original data:

```{r}
genotypesA <- c("AA", "Aa")
genotypesAF <- factor(genotypesA, levels = c("AA", "Aa", "aa"))
```
